---
title: "midterm_AK"
author: "Alex Knitter"
date: '2022-10-20'
output: html_document
---

```{r load-packages}
# load necessary packages
pacman::p_load(readr, tidyverse)
```


## a) importing the dataset

```{r load-data}
# import the dataset from github
titanic <- read_csv('https://raw.githubusercontent.com/kijohnson/Data/main/titanic.csv')
```


## b) determining the median age for by passenger class

```{r}
# get summary of age by passenger class (includes median and other descriptive stats)
by(titanic$Age, titanic$Pclass, summary)
```

**Passenger Class 1 Median Age:** 38.5 years

**Passenger Class 2 Median Age:** 29.0 years

**Passenger Class 3 Median Age:** 24.0 years


## c) a plot that shows medians, spread of data, and outliers for the continuous variable age according to passenger class

```{r}
# Recode Pclass to factor variable for plotting by class
titanic <- titanic %>%
  mutate(Pclass = recode_factor(Pclass,
                                "1" = "Class 1",
                                "2" = "Class 2",
                                "3" = "Class 3"))

# Chose boxplots to show medians, spread (quartiles), and outliers
# Generate boxplot of Age by Passenger Class
ggplot(data = titanic, aes(x = Pclass, y = Age)) +
  geom_boxplot() +
  # add labels for axes, title, and caption
  labs(x = "Passenger Class", 
       y = "Age",  
       title = "Age of Titanic Passengers by Class", 
       caption="Titanic data, Github.com/kijohnson")

# Export the above plot as an image for Midterm Question 3
ggsave(filename = "Q3_Boxplot.png",
       width = 4, height = 4, units = "in", bg="white")
```


## QUESTION 4 b. Next, add code to determine the percentage of males and females by class. Save this new version and push the new version to your Github repository.

```{r}
# creating table of % of each sex within each class
titanic %>%
  # group by the relevant variables, passenger class and sex
  group_by(Pclass, Sex) %>%
  # get count of each sex within each class
  summarise(n = n()) %>%
  # Calculate percents from counts
  # ie num of female passengers in Class 1 div by total passengers in Class 1
  mutate(Percent = n/sum(n)*100)
```

